---
title: "Tidlyr"
author: "MarieJ"
format: html
execute: 
  keep-md: true
  
---
execute: keep-md: create a md file 

## Tdlyr and ggplot2

```{r}
#| message:: false
here::i_am("gitr101.Rproj")
library(here)
library(ggplot2)
library(dplyr)
library(tidyr)
```
 not include the render in the git representory (ignore everything execpt the ignore fill and qmd file)
 
### Multiple lines on a single graph
 
```{r}
eustock <- as.data.frame(EuStockMarkets)
eustock <- eustock |> mutate(time=1:n())
```


```{r}
ggplot(eustock, aes(x=time,y=DAX)) +
  geom_line()
```
### Wrong solution
```{r}
ggplot(eustock, aes(x=time,y=DAX)) +
  geom_line() +
  geom_line(mapping= aes(y=CAC),color="red")

```
 
Problems: 
- y axis name is wrong
- colors must be specified manuallly
-we miss an association between colors and names 

### Tidlyr based solution 
```{r}
long_eustock <- eustock|> pivot_longer(-time, names_to = "Stock Index", values_to="Value")
```

```{r}
ggplot(long_eustock, aes(x=time,y=Value, group=`Stock Index`, color=`Stock Index`)) +
  geom_line()
```
```{r}
ggplot(long_eustock, aes(x=time,y=Value, color=`Stock Index`)) +
  geom_line()
```
=> same representation but may be harder for some calculation without the group 

```{r}
#| fig-height: 15
ggplot(long_eustock, aes(x=time,y=Value)) +
  geom_line() + 
  facet_wrap(~`Stock Index`, ncol=1, scale="free_y")
```

